FFT based convolution

### 2010-07-21 article/Article.tex
updated article
Ignore-this: edd347be772d143645f2e83076229454

darcs-hash:20100721124622-2fc9d-44b31b30d2baf85f62059a53e0069abad263685e.gz

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{Ga\"etan Lehmann}
% \authoraddress{}

@@

@@

@@

@@

@@

It also exposes the \verb$GreatestPrimeFactor$ accessors from \verb$itk::FFTPadImageFilter$ to let
the user choose between the extra performance added by the padding and the extra memory cost, and
\verb$PadMethod$ from the same filter to let the user choose the padding method: Zero flux Neumann
is used by default.

@@

Unlike most of my previous contributions, this work has {\em not} been supported by
the french National Institute for Agricultural Research (INRA) -- my employer.
I'd like to thank the INRA anyway, and more specifically the MIMA2 facility, who
provided the computer ressources used for this work.

@@

Some other work may be done on this basis. An non exhaustive list includes:
\begin{itemize}
 \item binary erosion/dilation filters based on FFT;
 \item a set of deconvolution filters;
 \item a more memory efficient implementation of the convolution filter using the Overlap-Save algorithm;
 \item allow the user to create a plan with \verb$FFTW_PATIENT$, \verb$FFTW_MEASURE$, or
       \verb$FFTW_EXHAUSTIVE$, so that FFTW can create better plan with \verb$FFTW_ESTIMATE$ in a
       later run;
 \item integrate FFTW's wisdom to improve FFT performance in ITK.
\end{itemize}

>>>>

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{Ga\"etan Lehmann}
\authoraddress{{$^1$}INRA, UMR 1198; ENVA; CNRS, FRE 2857, Biologie du
D\'eveloppement et 
Reproduction, Jouy en Josas, F-78350, France.}

@@

\subsection{Clipping}

Because of the numerical precision used in the FFT transforms, the values can sometime be greater or
smaller than the possible values of the output image.
The values are clipped to fit in the range of possible values of the output image.

@@

\subsubsection{Plan optimization}

The ITK classes have been modified in order to create optimized FFTW plan. Because the computation
of the optimized plan can take a lot of time, the default is still \verb$ESTIMATE$.

The optimization level can be changed globally only at this time, with the method
\verb$itk::FFTWLock::SetGlobalOptimizationLevel()$, or by setting the environment variable
\verb$ITK_FFTW_OPTIMIZATION_LEVEL$.

\subsubsection{Wisdom integration}

The optimized plans are automatically loaded and saved from/in the files \verb$~/itkwisdomf$ and \verb$~/itkwisdom$,
or in the file name stored in the environment variable \verb$ITK_FFTW_WISDOM _FILE_BASE_NAME$ if this
variable exists.
They are saved automatically when the program exits, and can also be saved with some dedicated methods
in \verb$itk::FFTWLock$.

\subsection{ClipImageFilter}

\verb$itk::ClipImageFilter$ is a filter very similar to \verb$itk::CastImageFilter$ excepted that it
take cares to clip the too small or to great values, instead of simply casting them in the output
pixel type.

As \verb$itk::CastImageFilter$, it is able to process zero pixel if the output and the input are of the
same type and the filter can be run in place.

@@

\subsection{FFTConvolutionImageFilterBase}

\verb$itk::FFTConvolutionImageFilter$ is the base class to use for all the FFT based convolution transforms. It
contains several methods designed to make easier the implementation of the FFT based convolution filters. There
are 
\begin{itemize}
 \item several kind of methods to prepare the input image and kernel;
 \item a method to register the progress of some additional filters;
 \item an easy access to the padded region and the oddity of the size on the x axis;
 \item several kind of methods to produce the output.
\end{itemize}

It is implemented as a subclass of \verb$itk::FFTPadImageFilter$ to expose the padding options and avoid
code duplication.

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.35]{inheritance}
\caption{New and modified classes inheritances.\label{fig:conv-pipeline}}
\end{figure}

@@

It is based on \verb$itk::FFTConvolutionImageFilter$.

@@

@@

@@

Some other work may be done on this basis. An non exhaustive list includes:
\begin{itemize}
 \item binary erosion/dilation filters based on FFT;
 \item a set of deconvolution filters;
 \item a more memory efficient implementation of the convolution filter using the Overlap-Save algorithm;
 \item integrate FFTW's wisdom to improve FFT performance in ITK.
\end{itemize}
